---
title: "adventr4"
author: "Nival Kolambage"
date: "5/27/2020"
output: html_document
---

Loading the data on Alice's relationship ratings of Zack over several weeks and creating a tidyverse tibble from it.:

```{r}
library(tidyverse)

ras_tib <- readr::read_csv("../data/ais_04_ras.csv")
```


Viewing the tibble :
```{r}
ras_tib
```


Finding the mean and median of Alice's ratings. Mean function structure *mean(variable, trim = 0, na.rm = FALSE)*. Median is the same except it has no trim because that wouldn’t make sense (the median is effectively the data with a 50% trim):

```{r}
mean(ras_tib$Rating, na.rm = TRUE)
median(ras_tib$Rating, na.rm = TRUE)

```


Zach and Milton spot an unusual score (an outlier) in Alice’s data by plotting a histogram. Let's explore that!

```{r}
ggplot2::ggplot(data = ras_tib, aes(Rating)) +
  geom_histogram(fill = 'dark green', alpha = 0.7, binwidth = 1) +
  labs(y = "Frequency", x = "Alice's relationship ratings(7-35)") +
  coord_cartesian(xlim = c(7, 35)) +
  scale_x_continuous(breaks = 7:35)
  theme_()
```


Filtering out the outlier '11'.

```{r}
ras_no_outlier <- ras_tib %>% 
  dplyr::filter(Rating != 11)
ras_no_outlier
```


Calculating the mean and median ratings from the filtered tibble :

```{r}
mean(ras_no_outlier$Rating, na.rm = TRUE)
median(ras_no_outlier$Rating, na.rm = TRUE)
```


Milton also calculates the variance, standard deviation and inter-quartile range (IQR) of Alice’s ratings (without the outlier). IQR also has an arg called 'type' which specifies the way to calculate IQR. default is 7. IQR function structure : *IQR(variable_name, na.rm = FALSE, type = 7)*

```{r}
var(ras_no_outlier$Rating, na.rm = TRUE)
sd(ras_no_outlier$Rating, na.rm = TRUE)
IQR(ras_no_outlier$Rating, na.rm = TRUE, type = 7)
```

Above calculations *with* the Outlier.

```{r}
print('Mean is:')
mean(ras_tib$Rating, na.rm = TRUE)
print('Median is:')
median(ras_tib$Rating, na.rm = TRUE)
print('Variance is:')
var(ras_tib$Rating, na.rm = TRUE)
print('Standard Deviation is:')
sd(ras_tib$Rating, na.rm = TRUE)
print('Interquartile range is:')
IQR(ras_tib$Rating, na.rm = TRUE, type = 8)
```
























