---
title: "adventr2"
author: "Nival Kolambage"
date: "5/26/2020"
output:
  pdf_document: default
  html_document: default
---

#This is the 'adventr' tutorial by Prof.Andy Field.

##Adventr_02

Names of the Band Members
```{r}
name <- c('Sudari', 'Nival', 'Nishali', 'Naveen')
name
```

Age of the members
```{r}
age <- c(27, 30, 19, 26)
```

Sex of members
```{r}
sex <- c(1, 0, 1, 0)
factor(sex, levels = 0:1, labels = c('Male', 'Female'))
sex
```
Alternative way to factorize sex: foracts pkg from tidyverse
```{r}
sex <- c("Female", "Male", "Female", "Male")
sex <- forcats::as_factor(sex)
sex
sex <- forcats::fct_relevel(sex, "Male", "Female")
sex
```

Loading the data that Prof Pincus uses to help Zach explore what attributes women look for in men in Chapter 3 of the book

```{r}
ha_tib <- readr::read_csv("../data/ais_ch03_ha.csv")
ha_tib
```

Accessing 'Ambitious' in ha_tib
```{r}
ha_tib$Ambitious
```


Squaring 'Humour'

```{r}

library(dplyr)
ha_tib <- ha_tib %>%
  dplyr::mutate(humour_sq = (Humour**2))
ha_tib
```

Filtering girls less than 16

```{r}
ha_tib_16 <- ha_tib %>%
  dplyr::filter (ageYears >= 16)
ha_tib_16
```
```
















